---
title: "K&D entropy analysis"
output: html_document
date: "2026-05-23"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}
# install.packages("slider")

library(tidyverse)
library(dplyr)
library(ggplot2)
library(slider)
library(lme4)
```

```{r load data}
df_someof <- read.csv("04_results_formatted.csv")
demo_someof <- read.csv("04_subject_info.csv")

df_some <- read.csv("05_results_formatted.csv")
demo_some <- read.csv("05_subject_info.csv")

df_some = df_some %>%
  mutate(workerid = workerid + 1200)

demo_some = demo_some %>%
  mutate(workerid = workerid + 1200)

df = rbind(df_someof, df_some)
demo = rbind(demo_someof,demo_some)

df = df %>%
  dplyr::select(-c(X))

df$logRT = log(df$rawRT)

df$rt = df$rawRT
```

### Before Exclusions
**Number of participants tested:**
```{r before_exclusion, echo=FALSE}
length(demo$workerid)
```
**Participants in each condition:**
```{r before_exclusion_2, echo=FALSE}
workers = df %>%
  filter(slide_type == "practice_trial")
table(workers$Answer.condition) /4
```

### Exclusions

**Non-unique participants (remove all attempts):**
```{r not_unique_exclude, echo=FALSE}
not_unique = demo$workerid[duplicated(demo$workerid)]
not_unique

#remove non-unique
df = df[!(df$workerid %in% not_unique),]
demo = demo[!(demo$workerid %in% not_unique),]
```

**Participants whose native language is not english:**
```{r lang_exclude, echo=FALSE}
lang_exclude = demo %>%
  dplyr::select(workerid, language) %>%
  unique() %>%
  filter(str_detect(language,regex("english", ignore_case=TRUE),negate=TRUE)) %>%
  filter(str_detect(language,regex("eng", ignore_case=TRUE),negate=TRUE)) %>%
  filter(str_detect(language,regex("egnlish", ignore_case=TRUE),negate=TRUE)) %>%
  filter(str_detect(language,regex("United States", ignore_case=TRUE),negate=TRUE)) %>%
  filter(str_detect(language,regex("en", ignore_case=TRUE),negate=TRUE)) %>%
  filter(str_detect(language,regex("ernglish", ignore_case=TRUE),negate=TRUE)) %>%
  filter(str_detect(language,regex("rnglish", ignore_case=TRUE),negate=TRUE)) %>%
  filter(str_detect(language,regex("USA", ignore_case=TRUE),negate=TRUE)) %>%
  filter(str_detect(language,regex("American", ignore_case=TRUE),negate=TRUE)) %>%
  filter(str_detect(language,regex("United Kingdom", ignore_case=TRUE),negate=TRUE)) %>%
  filter(str_detect(language,regex("ehglish", ignore_case=TRUE),negate=TRUE))

lang_exclude

lang_exclude = lang_exclude %>%
  dplyr::select(workerid)

df = df[!(df$workerid %in% lang_exclude$workerid),]
```
**Participants who got at least three practice trials wrong:**
```{r practice_exclude, echo=FALSE}
practice = df %>% 
  filter(str_detect(slide_type,"practice_trial")) %>%
  mutate(WrongAnswer = ((audio=="none.wav" & image=="13" & key=="Yes")|(audio=="all.wav" & image=="13" & key=="No")|(audio=="all.wav" & image=="0" & key=="Yes")| (audio=="none.wav" & image=="0" & key=="No"))) %>% 
  group_by(workerid) %>% 
  count(WrongAnswer) %>% 
  filter(WrongAnswer == TRUE & n>2)

#table(practice$n)

#df = df[!(df$workerid %in% practice$workerid),]
```
**Participants who got the audio check wrong more than one once:**
```{r audio_exclude, echo=FALSE}
audio_fail = df %>%
  filter(str_detect(slide_type,"audio_check")) %>%
  mutate(WrongAnswer = key != "Four") %>%
  group_by(workerid) %>% 
  count(WrongAnswer) %>%
  filter(WrongAnswer == TRUE & n>1)

#table(audio_fail$workerid)

#df = df[!(df$workerid %in% audio_fail$workerid ),]
```
**Participants who got the second comprehension question wrong more than twice:**
```{r comp_exclude, echo=FALSE}
comp1_fail = df[df$slide_type=="comprehension_check_1" & ((df$Answer.condition=="all_QUD" & df$key!="Empty") | (df$Answer.condition=="any_QUD" & df$key!="Jam")),]

comp2_fail = df[df$slide_type=="comprehension_check_2" & ((df$Answer.condition=="all_QUD" & df$key!="Empty") | (df$Answer.condition=="any_QUD" & df$key!="Jam")),]

comp1_wrong = comp1_fail %>%
  group_by(workerid,Answer.condition) %>%
  count() 

table(comp1_wrong$Answer.condition)

comp2_wrong = comp2_fail %>%
  group_by(workerid,Answer.condition) %>%
  count() 

table(comp2_wrong$Answer.condition)

#table(comp1_fail$workerid)
#table(comp2_fail$workerid)
table(comp1_fail$Answer.condition)
table(comp2_fail$Answer.condition)

comp2_fail =  comp2_fail %>%
  group_by(workerid) %>%
  count() %>%
  filter(n>2)

comp2_fail

df = df[!(df$workerid %in% comp2_fail$workerid ),]
```
**Participants with accuracy of lower than 85% on non-critical trials with "some","none","all" and numbers below 6:**
```{r accuracy_exclude, echo=FALSE}
yes_list = c("some2","some5","some8","some11",
             "all13",
             "none0",
             "zero0","two2","five5","eight8","eleven11","thirteen13")

#exclude_list = c("seven11","eight8","eight13","nine5","ten5","ten13","eleven0","eleven8","eleven11","twelve2","thirteen8")

exclude_list_1 = c("seven11","eight8","eight13","ten13","eleven8","eleven11","thirteen8")
exclude_list_2 = c("eight13","four5","seven11","two5","one11","ten13","five13","two8")

accuracy = df %>% 
  filter(str_detect(slide_type,"trial")) %>% 
  filter(str_detect(slide_type,"practice", negate = TRUE)) %>%
  mutate(concat = str_c(str_sub(audio,1,end=-5),image)) %>%
  filter(str_detect(concat,"some13", negate = TRUE)) %>%
  filter(!key=="Late") %>%
  filter(!concat %in% exclude_list_1) %>%
  filter(!concat %in% exclude_list_2) %>%
  mutate(rightAnswer = ifelse(concat %in% yes_list,"Yes","No")) %>%
  mutate(gaveRightAnswer = ifelse(key==rightAnswer,"1","0"))

answerNumber = accuracy %>%
  group_by(workerid) %>%
  count(workerid) %>%
  mutate(answerNm = n)

low_accuracy = accuracy %>%
  group_by(workerid,gaveRightAnswer) %>%
  count(gaveRightAnswer) %>%
  merge(answerNumber[ ,c("workerid","answerNm")], by="workerid",all.x=TRUE) %>%
  mutate(accuracy=ifelse(gaveRightAnswer=="1",n*100/answerNm,0)) %>%
  filter(gaveRightAnswer=="1") %>%
  filter(accuracy < 85)

low_accuracy

#latest exp
deneme1 = df %>%
  dplyr::select(workerid,Answer.condition) %>%
  unique()

deneme = low_accuracy %>%
  merge(deneme1[ ,c("workerid","Answer.condition")], by="workerid",all.x=TRUE)

table(deneme$Answer.condition)
#

df = df[!(df$workerid %in% low_accuracy$workerid),]
```
#### Additional Exclusions
**Participants who gave more than 5 very slow (logRT>20) responses:**
```{r slowworker_exclude, echo=FALSE}
slow_workers = df %>%
  filter(str_detect(slide_type,"trial")) %>% 
  filter(str_detect(slide_type,"practice", negate = TRUE)) %>% 
  mutate(slowResponse = logRT>30) %>%
  group_by(workerid) %>%
  count(slowResponse) %>%
  filter((slowResponse == TRUE)&(n>5))
slow_workers
  
df = df[!(df$workerid %in% slow_workers$workerid),]
```
**Responses that are faster than the onset of the quantifier (rawRT<600):**
```{r fast_trial_exclude, echo=FALSE}
df$ID <- seq.int(nrow(df))

fast = df %>%
  filter((slide_type=="trial_1"|slide_type=="trial_2") & rawRT<600)
nrow(fast)

fast = fast %>%
  dplyr::select(ID)

df = df[!(df$ID %in% fast$ID),]

```

**Responses that are very slow (logRT>20):**
```{r slow_trial_exclude, echo=FALSE}
slow = df %>%
  filter((slide_type=="trial_1"|slide_type=="trial_2") & logRT>20)
nrow(slow)

slow = slow %>%
  dplyr::select(ID)

df = df[!(df$ID %in% slow$ID),]

df = df %>%
  dplyr::select(-c(ID))
```
## After Exclusions
**Number of participants:**
```{r after_exclusions, echo=FALSE}
length(unique(df$workerid))
```
**Participants left in each condition:**
```{r after_exclusions_2, echo=FALSE}
workers = df %>%
  filter(slide_type == "practice_trial")
table(workers$Answer.condition) /4

```

```{r clean up}
dat <- df %>%
  filter(image == "13" & (audio == "some.wav" | audio == "summa.wav"),
         key != "Late",
         speaker_cond != "noQUD") %>%
  rename(QUD = speaker_cond) %>%
  mutate(partitive = ifelse(audio == "some.wav", "non-partitive", "partitive"))

dat <- dat %>%
  group_by(workerid) %>%
  mutate(consistent = if_else(n_distinct(key) == 1, 1, 0)) %>%
  ungroup()

dat <- dat %>%
  group_by(workerid) %>%
  mutate(type = ifelse(sum(key == "No") > sum(key == "Yes"), "pragmatic",
                       ifelse(sum(key == "No") < sum(key == "Yes"), "literal", "mixed"))) %>%
  mutate(nondominant_responses = ifelse(type == "pragmatic", sum(key == "Yes"),
                                     ifelse(type == "literal", sum(key == "No"), 4))) %>%
  arrange(slide_number, .by_group = TRUE) %>%
  mutate(order = row_number()) %>%
  ungroup()

count <- dat %>%
  group_by(partitive, QUD) %>%
  summarise(n = length(unique(workerid))) %>%
  ungroup()

```

```{r calculate entropy}
entropy <- function(p) {
  ifelse(p %in% c(0, 1), 0, -p * log2(p) - (1 - p) * log2(1 - p))
}

dat_summ <- dat %>%
  group_by(workerid) %>%
  arrange(workerid, order) %>%
  mutate(
    window = order - 2,
    implicature = slide_int(key, ~sum(.x == "No"), 
                            .before = 2, .after = 0, .complete = TRUE)
  ) %>%
  ungroup()
  
dat_summ <- dat_summ %>%
  filter(!is.na(implicature)) %>%
  mutate(p = implicature / 3,
         H = entropy(p))
```

```{r stats}
dat_summ$partitive <- factor(dat_summ$partitive)
dat_summ$QUD   <- factor(dat_summ$QUD)

contrasts(dat_summ$partitive) <- contr.sum(nlevels(dat_summ$partitive))
contrasts(dat_summ$QUD)   <- contr.sum(nlevels(dat_summ$QUD))

contrasts(dat_summ$partitive)
contrasts(dat_summ$QUD)

dat_summ$window_scaled <- scale(dat_summ$window)

m <- glmer(H ~ window_scaled*partitive*QUD + (1 + window_scaled | workerid),
          family = 'binomial',
          data = dat_summ)

summary(m)
```

```{r plot by QUD and partitivity}
dat_summ1 <- dat_summ %>%
  group_by(window, partitive, QUD) %>%
  summarize(entropy = mean(H),
            se = sd(H)/sqrt(n())) %>%
  ungroup()

base_size = 16

ggplot(dat_summ1, aes(x = window, y = entropy)) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymin = entropy - se, ymax = entropy + se), width = 0.2) +
  scale_x_continuous(breaks = unique(dat_summ1$window)) +
  labs(x = "window (width = 3)", y = "Shannon entropy") +
  facet_grid(partitive ~ QUD) +
  theme_bw(base_size = base_size) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(size = rel(1)))
```


```{r plot aggregated}
dat_summ2 <- dat_summ %>%
  group_by(window) %>%
  summarise(avg_entropy = mean(H),
            se = sd(H)/sqrt(n())) %>%
  ungroup()

ggplot(dat_summ2, aes(x = window, y = avg_entropy)) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymin = avg_entropy - se, ymax = avg_entropy + se), width = 0.2) +
  scale_x_continuous(breaks = unique(dat_summ1$window)) +
  labs(x = "window (width = 3)", y = "Shannon entropy") +
  theme_bw(base_size = base_size) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(size = rel(1)))
```
